Online Multiclass Linear Classification with Bandit Feedback

### 2018-11-28 arxiv-version/main.tex
Work on proof of auxiliary lemma

Before we state the theorems, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
$$
p(x) = p(x_1, x_2, \dots, x_d) = \sum_{i_1, i_2, \dots, i_d} c_{i_1, i_2, \dots, i_d} x_1^{i_1} x_2^{i_2} \dots x_d^{i_d}
$$
where the sum ranges over a finite set of tuples $(i_1, i_2, \dots, i_d)$ of
non-negative integers and $c_{i_1, i_2, \dots, i_d}$ is a real coefficient. The
\emph{degree} of a polynomial $p$, denoted by $\deg(p)$, is the largest value of
$i_1 + i_2 + \dots + i_d$ for which the coefficient is non-zero. The \emph{norm
of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{i_1, i_2, \dots, i_d} \left(c_{i_1, i_2, \dots, i_d} \right)^2 } \; .
$$

@@

\subsection{Proof of \autoref{theorem:polynomial-approximation-1}}
\label{section:proof-of-polynomial-approximation-1}
TODO

>>>>

Before we state the theorems, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
$$
p(x) = p(x_1, x_2, \dots, x_d) = \sum_{i_1, i_2, \dots, i_d} c_{i_1, i_2, \dots, i_d} x_1^{i_1} x_2^{i_2} \dots x_d^{i_d}
$$
where the sum ranges over a finite set of $d$-tuples $(i_1, i_2, \dots, i_d)$ of
non-negative integers and $c_{i_1, i_2, \dots, i_d}$ is a real coefficient. The
\emph{degree} of a polynomial $p$, denoted by $\deg(p)$, is the largest value of
$i_1 + i_2 + \dots + i_d$ for which the coefficient is non-zero. The \emph{norm
of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{i_1, i_2, \dots, i_d} \left(c_{i_1, i_2, \dots, i_d} \right)^2 } \; .
$$
It is not hard to see that this indeed a norm, since we can interpret it as
a Euclidean norm of the vector of the coefficients of the polynomial.

@@

\subsection{Proof of \autoref{theorem:polynomial-approximation-1}}
\label{section:proof-of-polynomial-approximation-1}

\begin{lemma}[Inequality]
\label{lemma:inequality}
For any real numbers $b_1, b_2, \dots, b_n$,
$$
\left( \sum_{i=1}^n b_i \right)^2 \le n \sum_{i=1}^n b_i^2 \; .
$$
\end{lemma}

\autoref{lemma:inequality} follows from Cauchy-Schwartz inequality applied to
vectors $(b_1, b_2, \dots, b_n)$ and $(1,1,\dots,1)$.

\begin{proposition}[Properties of the norm of polynomials]
\hspace{1cm} % Dummy space
\begin{enumerate}
\item Let $p_1, p_2, \dots, p_n$ be multivariate polynomials and
let $p(x) = \prod_{j=1}^n p_j(x)$ be their product. Then, $\norm{p}^2 \le n^{\sum_{j=1}^n \deg(p_j)} \prod_{j=1}^n \norm{p_j}^2$.
\item Let $q$ be a multivariate polynomial of degree at most $s$ and let $p(x) = (q(x))^n$. Then, $\norm{p}^2 \le n^{ns} \norm{q}^{2n}$.
\item Let be $p_1, p_2, \dots, p_n$ be multivariate polynomials. Then, $\norm{\sum_{j=1}^n p_j}^2 \le n \sum_{j=1}^n \norm{p_j}^2$.
\end{enumerate}
\end{proposition}

\begin{proof}
Using multi-index notation we can write any multivariate polynomial $p$ as
$$
p(x) = \sum_A c_A x^A
$$
where $A = (i_1, i_2, \dots, i_d)$ is a multi-index (i.e. a $d$-tuple of
non-negative integers), $x^A = x_1^{i_1} x_2^{i_2} \dots x_d^{i_d}$ is a
monomial and $c_A = c_{i_1, i_2, \dots, i_d}$ is the corresponding real
coefficient. The sum is over a finite subset of $d$-tuples of non-negative
integers. Using this notation, the norm of a polynomial $p$ can be written as
$$
\norm{p} = \sqrt{\sum_A (c_A)^2} \; .
$$
For a multi-index $A = (i_1, i_2, \dots, i_d)$ we define its $1$-norm as
$\norm{A}_1 = i_1 + i_2 + \dots + i_d$.

To prove the first part, we express $p_j$ as
$$
p_j(x) = \sum_{A_j} c^{(j)}_{A_j} x^{A_j} \; .
$$
If $p$ has expansion $p(x) = \sum_A c_A x^A$ then its coefficient $c_A$ for any multi-index $A$ is
$$
c_A = \sum_{\substack{(A_1, A_2, \dots, A_n) \\ A_1 + A_2 + \dots + A_n = A}} c^{(1)}_{A_1} c^{(2)}_{A_2} \cdots c^{(n)}_{A_n} \; .
$$
Therefore,
\begin{multline*}
\norm{p}^2
= \sum_{A} (c_A)^2
= \sum_{A} \left( \sum_{\substack{(A_1, A_2, \dots, A_n) \\ A_1 + A_2 + \dots + A_n = A}} c^{(1)}_{A_1} c^{(2)}_{A_2} \cdots c^{(n)}_{A_n} \right)^2
= \sum_{A} \left( \sum_{\substack{(A_1, A_2, \dots, A_n) \\ A_1 + A_2 + \dots + A_n = A}} \prod_{j=1}^n c^{(j)}_{A_j} \right)^2
\end{multline*}
and
\begin{multline*}
\prod_{i=1}^n \norm{p_i}^2
= \prod_{i=1}^n \left( \sum_{A_i} (c^{(i)}_{A_i})^2 \right)
= \sum_{(A_1, A_2, \dots, A_n)} \prod_{j=1}^n (c^{(j)}_{A_j})^2
= \sum_{(A_1, A_2, \dots, A_n)} \left( \prod_{j=1}^n c^{(j)}_{A_j} \right)^2 \\
= \sum_A \sum_{\substack{(A_1, A_2, \dots, A_n) \\ A_1 + A_2 + \dots + A_n = A}} \left( \prod_{j=1}^n c^{(j)}_{A_j} \right)^2
\end{multline*}
where in both cases the outer sum is over multi-indices $A$ such that $\norm{A}_1 \le \deg(p)$.
To finish the proof, it is sufficient to prove that for any $A$ such that $\norm{A}_1 \le \deg(p)$,
$$
\left( \sum_{\substack{(A_1, A_2, \dots, A_n) \\ A_1 + A_2 + \dots + A_n = A}} \prod_{j=1}^n c^{(j)}_{A_j} \right)^2
\le n^{\deg(p)} \sum_{\substack{(A_1, A_2, \dots, A_n) \\ A_1 + A_2 + \dots + A_n = A}} \left( \prod_{j=1}^n c^{(j)}_{A_j} \right)^2 \; .
$$
This inequality follows from \autoref{lemma:inequality} as long we can show that
the number of $n$-tuples $(A_1, A_2, \dots, A_n)$ such that $A_1 + A_2 + \dots +
A_n = A$ is upper bounded by $n^{\deg(p)}$.

The second part follows from the first part by setting $p_1 = p_2 = \dots p_n = q$.

To prove the third part we use generalized triangle inequality and an
\autoref{lemma:inequality}. We have
\begin{align*}
\norm{\sum_{j=1}^n p_j}^2 = \left( \norm{\sum_{j=1}^n p_j} \right)^2 \le \left(\sum_{j=1}^n \norm{p_j} \right)^2 \le n \sum_{j=1}^n \norm{p_j}^2 \; .
\end{align*}
\end{proof}
